---
title: "HW5"
author: "Yuelin Long"
date: "March 10, 2019"
output: html_document
---


```{r}
#Load libraries
library(tidyverse)
library(pscl)
```


##Problem 1

####Question 1
```{r}
#Load raw data
crab_data = read.delim("HW5-crab.txt", sep = "") %>% 
  select(-number)

m1.fit = glm(Sa~W, family=poisson, data=crab_data)
summary(m1.fit)

#Goodness of fit
m1.fit$deviance
n = nrow(crab_data)
p1 = 2
df = n-p1
qchisq(0.05,df)
pval=1-pchisq(m1.fit$deviance,df=df)
pval
```

Intepretation:

+ Intercept b0: Since the width of carapace cannot be 0, the intercept has no actual meaning.
+ Coefficient b1: With 1 unit change in carapace width, the log odds rate of the number of satellites of female crab would increase by 0.16405.


Goodness of fit:
The deviance is far greater than critical value. Therefore, reject the null hypothesis and conclude that the fit is bad. The small p value also confirms it.


#Question 2
```{r}
m2.fit = glm(Sa~W+Wt, family=poisson, data=crab_data)
summary(m2.fit)

#Goodness of fit
m2.fit$deviance
p2 = 3
df = n-p2
qchisq(0.05,df)
pval=1-pchisq(m2.fit$deviance,df=df)
pval
```

Intepretation:

+ Intercept b0: Since the width of carapace cannot be 0, nor can the weight be 0, the intercept has no actual meaning.
+ Coefficient of W (b1): Given weight fixed, with 1 unit change in carapace width, the log odds rate of the number of satellites of female crab would increase by 0.04590.
+ Coefficient of Wt (b2): Given catapace width fixed, with 1 unit change in weight, the log odds rate of the number of satellites of female crab would increase by 0.44744.


```{r}
#Estimate over-dispersion parameters
res.p2=residuals(m2.fit,type='pearson',data=crab_data) 
G2=sum(res.p2^2)
phi=G2/df

#Check whether over dispersion exist - Half normal plot
plot(qnorm((n+1:n+0.5)/(2*n+1.125)),sort(abs(res.p2)),xlab='Expected Half-Normal Order Stats',ylab='Ordered Abs Pearson Residuals')
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2) 
```

The data fit the dotted line better, and the slope of dotted line is quite different from the solid line. This indicates that there is over dispersion.

```{r}
#Fit model adjusted for over dispersion
summary(m2.fit,dispersion=phi)
```


Intepretation (same as unadjusted):

+ Intercept b0: Since the width of carapace cannot be 0, nor can the weight be 0, the intercept has no actual meaning.
+ Coefficient of W (b1): Given weight fixed, with 1 unit change in carapace width, the log odds rate of the number of satellites of female crab would increase by 0.04590.
+ Coefficient of Wt (b2): Given catapace width fixed, with 1 unit change in weight, the log odds rate of the number of satellites of female crab would increase by 0.44744.



##Problem 2
####Question 1
```{r}
#Load raw data
para_data = read.delim("HW5-parasite.txt", sep = "") %>% 
  select(-Sample, -starts_with("omit")) %>% 
  na.omit() %>% 
  dplyr::mutate(Area = factor(Area, levels = rep(1:4)),
         Year = factor(Year, levels = rep(1999:2001)))

m1.fit = glm(Intensity~., family=poisson, data=para_data)
summary(m1.fit)
```

Intepretation:

+ Intercept: Since the length of the fish can't be 0, the intercept has no actual meaning.
+ Coefficient of Year 2000: Given length of fish and area fixed, the log odds rate of the increase of number of parasites is 0.6702801 for Year 2000 verse year 1999.
+ Coefficient of Year 2001: Given length of fish and area fixed, the log odds rate of the increase of number of parasites is -0.2181393 for Year 2001 verse year 1999.
+ Coefficient of Length: Given year and area fixed, every 1 unit increase in the length of fish would increase the log odds rate of the number of parasites by -0.0284228.
+ Coefficient of Area2: Given year and length of fish fixed, the log odds rate of the increase of number of parasite by -0.2119557 in area 2 versus area 1.
+ Coefficient of Area3: Given year and length of fish fixed, the log odds rate of the increase of number of parasite by -0.1168602 in area 3 versus area 1.
+ Coefficient of Area4: Given year and length of fish fixed, the log odds rate of the increase of number of parasite by 1.4049366 in area 4 versus area 1.





####Question 2
```{r}
#Goodness of fit
m1.fit$deviance
n = as.numeric(nrow(para_data))
p1 = ncol(para_data) - 1
df = n-p1
qchisq(0.05,df)
pval=1-pchisq(m1.fit$deviance,df=df)
pval
```

Goodness of fit:
The deviance is far greater than critical value. Therefore, reject the null hypothesis and conclude that the fit is bad. The small p value also confirms it.

####Question 3
```{r}
m2.fit = zeroinfl(Intensity ~ Year + Length + Area | Year + Length + Area, data = para_data)
summary(m2.fit)
```

Intepretation:

For Poisson model:

Intepretation:

+ Intercept: Since the length of the fish can't be 0, the intercept has no actual meaning.
+ Coefficient of Year 2000: Given length of fish and area fixed, the log odds rate of the increase of number of parasites is 0.3919831 for Year 2000 verse year 1999.
+ Coefficient of Year 2001: Given length of fish and area fixed, the log odds rate of the increase of number of parasites is -0.0448455 for Year 2001 verse year 1999.
+ Coefficient of Length: Given year and area fixed, every 1 unit increase in the length of fish would increase the log odds rate of the number of parasites by  -0.0368067.
+ Coefficient of Area2: Given year and length of fish fixed, the log odds rate of the increase of number of parasite by 0.2687835 in area 2 versus area 1.
+ Coefficient of Area3: Given year and length of fish fixed, the log odds rate of the increase of number of parasite by 0.1463173 in area 3 versus area 1.
+ Coefficient of Area4: Given year and length of fish fixed, the log odds rate of the increase of number of parasite by 0.9448068 in area 4 versus area 1.


For Binomial model:

+ Intercept b0: Since the length of the fish can't be 0, the intercept has no actual meaning.
+ Coefficient of Year 2000: Given length of fish and area fixed, the log odds ratio of susceptible to parasites over not susceptible to parasites is -0.752119 for Year 2000 verse year 1999.
+ Coefficient of Year 2001: Given length of fish and area fixed, the log odds ratio of susceptible to parasites over not susceptible to parasites is 0.456535 for Year 2001 verse year 1999.
+ Coefficient of Length: Given year and area fixed, for every 1 unit change in length of the fish, the log odds ratio of susceptible to parasites over not susceptible to parasites would be -0.009889.
+ Coefficient of Area2: Given year and length of fish fixed, the log odds ratio of susceptible to parasites over not susceptible to parasites would be 0.718676 in area 2 versus area 1.
+ Coefficient of Area3: Given year and length of fish fixed, the log odds ratio of susceptible to parasites over not susceptible to parasites would be 0.657708 in area 3 versus area 1.
+ Coefficient of Area4: Given year and length of fish fixed, the log odds ratio of susceptible to parasites over not susceptible to parasites would be -1.022868 in area 4 versus area 1.


